---
title: "Lab Sesion 1"
author: "Giulia Bertoldo"
date: "2/22/2022"
output:
   html_document:
    theme: cosmo
    highlight: monochrome
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r}
library(haven)
library(dplyr)
library(corrplot)
```

# 1. General introduction to the dataset

The European Social Survey is a survey conducted across Europe every two years via face-to-face interviews. The survey measures "attitude, beliefs and behavior patterns" on a representative sample of people aged 15 and over resident within private households. Here we consider the 4th edition, run in 2008 in Belgium. 

## More information

* [Data documentation](https://www.europeansocialsurvey.org/docs/round4/survey/ESS4_data_documentation_report_e05_5.pdf).

* [Source questionnaire](https://www.europeansocialsurvey.org/docs/round4/fieldwork/source/ESS4_source_main_questionnaire.pdf)

* [Data](http://nesstar.ess.nsd.uib.no/webview/)

\newpage

# 2. Selection of variables for all the laboratories 

We use a subset of 21 variables. 

*D15-17 & D18-20 & & D21 - 26 & D27 - 29 & D1 - 6: Part of the block D1-50: Welfare includes attitudes towards welfare provision, size of claimant groups, views on taxation, attitutdes towards service delivery and likely future dependence on welfare.*

People have different views on what the responsibilities of governments should or should not be. For each of the tasks I read out please tell me on a score of 0-10 how much responsibility you think governments should have. 0 means it should not be governments? responsibility at all and 10 means it should be entirely governments? responsibility. Firstly to? (0 Not governments’ responsibility at all - 10 Entirely governments’ responsibility)

**Welfare Support** items:

* gvslvol Standard of living for the old, governments’ responsibility  
    +  ...**ensure a reasonable standard of living for the old?**
    + D15 - 17
    
* gvhlthc Health care for the sick, governments’ responsibility 
    + ...**ensure adequate health care for the sick?**
    + D15 - 17
  
* gvjbevn Job for everyone, governments’ responsibility 
    + ...**ensure a job for everyone who wants one?**
    + D15 - 17

* gvcldcr Child care services for working parents, governments’ responsibility 
    + ...**ensure sufficient child care services for working parents?**
    + D18-20 

* ggvslvue Standard of living for the unemployed, governments’ responsibility
    + ...**ensure a reasonable standard of living for the unemployed?**
    + D18-20
  
* gvpdlwk Paid leave from work to care for sick family, governments’ responsibility 
    + And how much responsibility do you think governments should have to... ...**provide paid leave from work for people who temporarily have to care for sick family members?**
    + D18-20 

**Economic Criticism** items: 

(D21-26) Using this card please tell me to what extent you agree or disagree that social benefits and services in [country]... (1 Agree strongly - 5 Disagree strongly) 

* sbstrec Social benefits/services place too great strain on economy
    + ....**place too great a strain on the economy?**
    + D21 - 26

* sbbsntx Social benefits/services cost businesses too much in taxes/charges  
    + ....**cost businesses too much in taxes and charges?**
    + D21 - 26

**Social Criticism** items: 

* sbprvpv Social benefits/services prevent widespread poverty 
    + ....**prevent widespread poverty?**
    + D21 - 26

* sbeqsoc Social benefits/services lead to a more equal society 
    + ....**lead to a more equal society?**
    + D21 - 26
    
* sbcwkfm Social benefits/services make it easier to combine work and family 
    + ....**make it easier for people to combine work and family life?**
    + D21 - 26



(D27-29) And to what extent do you agree or disagree that social benefits and services in [country]...
 
**Moral Criticism** items: 

* sblazy Social benefits/services make people lazy 
    + ...**make people lazy?**
    + D27 - 29

* sblwcoa Social benefits/services make people less willing care for one another
    + ...**make people less willing to care for one another?**
    + D27 - 29
    
* sblwlka Social benefits/services make people less willing look after themselves/family 
    + ....**make people less willing to look after themselves and their family?**
    + D27 - 29

**Other** items: 

(D1 - 6) Using this card, please say how much you agree or disagree with each of the following statements (1 Agree strongly - 5 Disagree strongly).
    
* dfincac Large differences in income acceptable to reward talents and efforts
    + **Large differences in people's incomes are acceptable to properly reward differences in talents and efforts**. 
    + D1 - 6

* smdfslv For fair society, differences in standard of living should be small
    + **For a society to be fair, differences in people's standard of living should be small.**
    + D1 - 6


*B30 -33: Part of the block D1-50: Politics, including: political interest, efficacy, trust, electoral and other forms of participation, party allegiance, socio-political orietnations.*  

Using this card, please say to what extent you agree or disagree with each of the following statements (1 Agree strongly - 5 Disagree strongly).

* gincdif Government should reduce differences in income levels
    + **The government should take measures to reduce differences in income levels**


* agea Respondent’s age

* eduyrs Years of full-time education completed

* gndr Gender (1 Male, 2 Female)

* hinctnta Household’s total net income, all sources (Deciles of the actual household income range in the given country.)

# 3. Variables used in laboratory session 1

**Welfare Support** items:

* gvslvol Standard of living for the old, governments’ responsibility  
    
* gvhlthc Health care for the sick, governments’ responsibility 
  
* gvjbevn Job for everyone, governments’ responsibility 

* gvcldcr Child care services for working parents, governments’ responsibility 

* ggvslvue Standard of living for the unemployed, governments’ responsibility
  
* gvpdlwk Paid leave from work to care for sick family, governments’ responsibility 

\newpage

# 2. Analysis plan

\newpage 

# 3. EFA

```{r}
# Import data
ess_df <- haven::read_sav("data/ESS4_belgium.sav")

# EFA
str(ess_df)
View(ess_df)
head(ess_df)

# number of subjects
nrow(ess_df)

# number of variables
ncol(ess_df)

# names of variables
names(ess_df)

# In this first lab, we are exploring two concepts:
# welfare support and welfare criticism. Let’s take a closer look at our items

# Select only variables of interest
ess_df_selected <- ess_df %>% select(
  ## Welfare support items ##
  gvslvol, # the old
  gvslvue, # the unemployed
  gvhlthc, # the sick
  gvcldcr, # working parents
  gvjbevn, # job for everyone
  gvpdlwk, # paid sick leave
  ##    Economic criticism items ##
  sbstrec, # strain on economy
  sbbsntx, # too much taxes
  ##    Social criticism items ##
  sbprvpv, # poverty
  sbeqsoc, # more equal society
  sbcwkfm, # work and family
  ##    Moral criticism items ##
  sblazy,  # people lazy
  sblwcoa, # care for others
  sblwlka  # look after others
)

# Descriptives
descriptive_ess <- as.data.frame(psych::describe(ess_df_selected))
(descriptive_ess <- dplyr::select(descriptive_ess,
                                  n,
                                  mean,
                                  sd,
                                  median,
                                  min,
                                  max,
                                  skew,
                                  kurtosis))

# Calculate variance-covariance matrix
# Welfare support items only
ess_df_welfare_supp <- ess_df %>% select(
  ## Welfare support items ##
  gvslvol, # the old
  gvslvue, # the unemployed
  gvhlthc, # the sick
  gvcldcr, # working parents
  gvjbevn, # job for everyone
  gvpdlwk  # paid sick leave
)
# Sample implied covariance matrix
(welfare_supp_cov <- cov(ess_df_welfare_supp,          # data frame
                        use = "pairwise.complete.obs" # remove NAs
))
# Visualize
(welfare_supp_cor <- cov2cor(welfare_supp_cov))
corrplot::corrplot(welfare_supp_cor,
                   is.corr = FALSE,       # whether is a correlation matrix
                   method = "circle",     # magnitude of covariances as circles
                   type = "upper",        # remove the bottom of the covariance matrix
                   addCoef.col = "black"  # add to the plot the coefficients
)

```





